Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-10-03 paper.tex
some restructuring
completed reproducibility

             Containeri
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        Reproducibility is a vital \tristan{critical?} component to all scientific workflows. 
        Without reproducibility, correctness of results could never be 
        determined. Reproducibility is defined as the ability to obtain similar 
        results by 
        recreating or executing the original workflow on the same or different
        infrastructure\note{cite Peng 2011}. It has been found that more than 70\% of
        scientific experiments are not reproducible \note{cite Mbaker}

@@

        This chapter will examine how different workflow engines tackle the 
        issue of reproducibility. Section~\ref{prov} will discuss how the 
        engines assimilate and distribute the provenance data. 
        Section~\ref{repcont} will examine how these engines use containers to
        facilitate reproducibility of the experiments.
        \section{Provenance Tracking}\label{prov}
            Provenance tracking is \weird{of vital importance} \tristan{twice} to scientific reporting.
            On its own, source code is insufficient to reproduce the workflow 
            described in a research paper. This is because source code does 
            not specify the specific dataset used nor the specific versions of
            tools and the environments that they were executed on. Minor 
            changes to any of these features can significantly affect results.

@@

            To to the importance of provenance tracking for scientific 
            workflows several workflow engines adapted strategies to capture 
            comprehensive provenance data. These include Nipype, PSOM, FastR, 
            LONI, Taverna, Pegasus, Nextflow, Galaxy. Other workflow engines,
            such as SPM, AA, Pydpiper,

        \section{Containerization}\label{repcont}
    \chapter{Discussion/Conclusion}

>>>>

             Containeri
            use of containers increase portability. none of the scientific 
            pipelines can be containerized and communicate with the scheduler 
            outside of the container
            big data frameworks can
    \chapter{Reproducibility}\label{reproducibility}
        Reproducibility is a critical component to all scientific workflows. 
        Without reproducibility, correctness of results could never be 
        determined. Reproducibility is defined as the ability to obtain similar 
        results by 
        recreating or executing the original workflow on the same or different
        infrastructure\note{cite Peng 2011}. It has been found that more than 70\% of
        scientific experiments are not reproducible \note{cite Mbaker}

@@

        This chapter will examine how different workflow engines tackle the 
        issue of reproducibility. Section~\ref{metadata} will investigate three 
        different provenance metadata specifications available (W3C PROV, NIDM,
        and BIDS Derivates).Section~\ref{prov} will discuss how the 
        engines assimilate and distribute the provenance data. 
        Section~\ref{repcont} will examine how these engines use containers to
        facilitate reproducibility of the experiments.
        \section{Metadata formats}\label{metadata}
            The recent increase in data-sharing resulted in the rise of BigData 
            in neuroscience. To continue the promotion of data sharing in 
            neuroscience, the derived data's metadata must be well understood
            by the community. Even if the provenance tracking is comprehensive,
            it is useless if it is only understood by some. To address this 
            issue, specifications for the standardization of metadata have been 
            proposed. Here we look at three provenance metadata specifications.
            The first is W3 PROV, a general-purpose provenance specification; 
            NIDM-results
            a comprehensive neuroimaging provenance specification; and BIDS 
            Derivatives the recent provenance addition to the family of 
            popular neuroimaging metadata specifications, BIDS.

            \subsection{W3C PROV}
            The W3C PROV specification came out from the need to explain the 
            origin of data shared on the internet. 
            The aim of the W3 PROV specification was gather information on
            the activities, entities and involved parties in the creation of 
            the data. This metadata would allow for assements of the quality, 
            reliability and trustworthiness to be performed. Sharing of the 
            provenance is faciliated through the use of widely-used formats 
            such as RDF and XML.

            The PROV model has three main objects involved: 1) entities, 2) 
            activities and 3) agents. Any kind of thing is considered an entity
            . Actions act upon entities to generate new ones. Activities have 
            agents associated to them. These agents are responsible for the 
            activity taking place. Agents can be individuals, companies and 
            software, amongst others.

            \subsection{NIDM-Results}
                NIDM-results was designed to improve the limited XML-based 
                Clinical and Experimental Data Exchange (XCEDE) neuroimaging
                provenance metadata format. It is a domain-specific extension
                to the W3 PROV model that aims to provide comprehensive and
                machine-readable provenance from raw data to its final stage.
                However, it is currently limited in the provenance 
                information it can provide.

            \subsection{BIDS Derivatives}
                The BIDS-Derivatives metadata specification is the newest 
                addition to the BIDS family of neuroimaging-related 
                specifications. It is the sister
                specification to the widely adopted BIDSRaw metadata format 
                and it remains compatible with the BIDS Raw specification. 
                Unlike NIDM-Results, BIDS derivatives, like BIDS Raw, do not
                use the W3C PROV specification. Rather, the provenance can be
                derived from its folder structure and filename. Any additional
                metadata can be found in a JSON sidecar file.

        \section{Provenance Tracking}\label{prov}
            Provenance tracking is crucial to scientific reporting.
            On its own, source code is insufficient to reproduce the workflow 
            described in a research paper. This is because source code does 
            not specify the specific dataset used nor the specific versions of
            tools and the environments that they were executed on. Minor 
            changes to any of these features can significantly affect results.

@@

            To address the importance of provenance tracking for scientific 
            workflows several workflow engines adapted strategies to capture 
            provenance metadata. Some leverage the W3C PROV specification (e.g.
            Nipype, FastR, Taverna), while others may chose to use NIDM (e.g. 
            SPM). The 
            remainder use a custom format to share their metadata.

            Although Nipype, FastR and Taverna all capture provenance using 
            the W3C Prov specification, they differ slightly in how they
            present provenance
            information and what they capture.
            Currently, Nipype's support for W3C PROV is experimental. However,
            it is capable of outputting provenance using the W3C PROV data 
            model for each workflow interface, as well as the workflow as a 
            whole.

            FastR performs provenance very similarly to Nipype. However,
            FastR's agents can be Fastr, Networks, Tools, Nodes. Additionally,
            activities, in
            FastR map to Jobs and entities map to the data.

            Taverna differs from Nipype and Taverna in that it provides 
            tracking of
            both workflow runs and of the derived data.

            Likewise to the engines that capture provenance that follows the
            W3C PROV specification, all MATLAB-based workflow engines capture
            provenance data similarly.
            By default, SPM saves an execution by writing the  workflow, with
            all its input parameters as a MATLAB file. Saving the workflow as
            a MATLAB file allows for distribution of the workflow and 
            reexecution. To export to NIDM-Results, the option must be 
            sslected.

            As in SPM's MATLAB workflow execution files, PSOM and AA 
            also save
            their workflow to a MATLAB file for future reexecution. 
            In addition, PSOM generates
            also files containing additional provenance information such 
            as username, date, time and operating system used.

            While all workflow engines capture generally the same type of 
            provenance information, they differ in how much provenance 
            information is captured and how it's done. LONI, Pegasus, Pydpiper 
            and Galaxy
            capture an above minimum amount of provenance data, whereas engines 
            like Nextflow, Make, Toil and MapReduce capture very limited  
            provenance 
            information. OpenMOLE handles provenance through encapsulation of
            tasks using CARE to simulate runtime environments
            and BigData frameworks like Spark and Dask mainly rely on 
            lineage.

            An above minimum amount of provenance details captured includes 
            capturing details on the workflow stages as well as the inputs and
            outputs. LONI stores this type of information in an XML file.

            Pegasus uses an SQL database to store its provenace. Data collected 
            includes runtime information,
            in addition to, data locations, software used and parameters 
            applied to the software.

            Pydpiper has no formal form of provenance tracking, but will save 
            essential pipeline information after each completed stage to a 
            serialized Python object. Data captured includes: the DAG, the 
            array of pipeline stages, the current stage counter, hash of stages 
            and output files and an array of stage statuses.

            As provenance is a major component of the Galaxy project, it 
            provides comprehensive provenance tracking. This includes gathering
            of metadata similarly to other workflow engines, but also allows
            users to add annotations to make the intent of their analysis known
            .

            The workflows that capture limited provenance information typically
            restrict themselves to capturing only runtime information.
            Nextflow generates an
            HTML report that contains details on execution commands, runtime, 
            CPU usage and status.

            Toil captures at most tool execution logs and provides runtime 
            statistics.

            Make is another workflow engine that does not provide built-in 
            provenance tracking. However, Make can be adapted to capture 
            provenance as in \note{cite make paper}.

            Similarly, MapReduce also captures runtime information and relays 
            it through a web UI.

            OpenMOLE tackles provanence differently by wrapping workflow tools
            into CARETasks. CARETasks simulate the runtime environment of the 
            execution, such that workflows can be executed on any Linux 
            distribution avoiding any silent errors that may arise from 
            executing a workflow in a different environment. An issue with 
            restricting provenance to CARETasks, however, is that provenance is
            a bit more than just simulating an environment. Provenance 
            information should also be accessible. Understanding of CARE should
            not be a prerequisite to obtaining provenance information.

            Dask and Spark rely on their lineage to relay details on which 
            tasks modified the data. They also provide a web UI that provides
            additional information on runtime statistics. Methods to improve 
            provenance tracking in Spark, particularly interactively are 
            currently being explored \note{cite bigdebug}. Unfortunately, both
            of them lack the ability to provide provenance capture of external
            tools at present.

        \section{Containerization}\label{repcont}

            To handle issues with numerical instability of tools and relieve 
            users from
            having to configure their environment to be almost identical to 
            that of the study they are trying to replicate, significant 
            emphasis has been placed on containerization. Similarly to 
            OpenMOLE's CARETasks, containerization provides an encapsulated
            environment for the tools to run which mimics the environment of 
            the original study. This ensures that users executing their 
            workflows in the same environments will obtain the same results.
            A major issue with containerization is that despite it controlling
            environments and permitting experiments to be reproducible, it does
            not solve the issue of numerical instability. Moreover, this 
            mentality may lead researchers to limit themselves to older tool 
            versions and further limit scientific advancement.

    \chapter{Discussion/Conclusion}
